The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-22 manuscript.tex

% !TeX root = RJwrapper.tex
\title{Containers and R---the Rockerverse and beyond}
\author{by Daniel Nüst, Dom Bennett, Robrecht Cannoodt, Dirk Eddelbuettel, Mark Edmondson, Colin Fay, Ellis Hughes, Sean Lopp, Ben Marwick, Heather Nolis, Jacqueline Nolis, Hong Ooi, Karthik Ram, Noam Ross, Lori Shepherd, Nitesh Turaga, Nan Xiao}

@@

\begin{figure}
\centering
\includegraphics{/home/daniel/git/o2r/rockerverse-paper/rockerverse.png}
\caption{Rockerverse hex sticker}
\end{figure}

@@

\textbf{Renjin} is an interpreter for the R language running on top of
the \href{https://en.wikipedia.org/wiki/Java_virtual_machine}{Java
Virtual Machine} (JVM) providing full two-way access between Java and R
code \citep{wikipedia_renjin_2018}. It was developed to combine the
benefits of R, such scripting and extension packages, with the JVM's
advantages in the areas of security, cross-platform availability, and
established position in enterprise settings. R extension packages need
to be specially compiled and are distributed via the Java package
manager \href{https://en.wikipedia.org/wiki/Apache_Maven}{Apache Maven},
cf.~\url{http://packages.renjin.org/packages} for available packages.
Packages are loaded on demand, i.e., at the first call to
\texttt{library()}. Not all R packages, especially one linking to binary
libraries, are available, e.g., \texttt{rgdal}
\footnote{\href{http://packages.renjin.org/package/org.renjin.cran/rgdal/1.4-4/build/1}{http://packages.renjin.org/package/org.renjin.cran/rgdal/1.4-4/build/1}}.
There are no offical Docker images for Renjin, but community-maintained
images for selected releases only are available under
\texttt{nuest/renjin} on Docker~Hub and GitHub at
\url{https://hub.docker.com/r/nuest/renjin} and
\url{https://github.com/nuest/renjin-docker} respectively. These images
expose the command line interface of Renjin in a similar fashion as
Rocker images and allow an easy evaluation of Renjin's suitability, but
are not intended for production use.

@@

\textbf{FastR} (\url{https://github.com/oracle/fastr}) is ``A
high-performance implementation of the R programming language, built on
GraalVM'' \citep{oracle_labs_oraclefastr_2020}. It is developed by
Oracle, connects R to the GraalVM ecosystem
\citep{wikipedia_graalvm_2019}, and also claims superior performance but
also targets full compatibility with base R
\citep{oracle_labs_oraclefastr_2020}. There are no offical Docker images
provided, but indepedently maintained experimental images and
\texttt{Dockerfile}s are provided by
\href{https://github.com/nuest/fastr-docker}{nuest/fastr-docker},
e.g.~\href{https://cloud.docker.com/repository/docker/nuest/fastr/}{\texttt{nuest/fastr}
on Docker Hub}.

@@

The \href{https://github.com/jupyter/docker-stacks/}{\textbf{Jupyter
Docker Stacks} project} are a set of ready-to-run Docker images
containing Jupyter applications and interactive computing tools
\citep{project_jupyter_jupyter_2018}. The \texttt{jupyter/r-notebook}
image includes R and ``popular packages'', and R is also included in the
catchall \texttt{datascience-notebook}
image\footnote{\href{https://jupyter-docker-stacks.readthedocs.io/en/latest/using/selecting.html}{https://jupyter-docker-stacks.readthedocs.io/en/latest/using/selecting.html}}.
For example, these images allow users to quickly start a Jupyter
Notebook server locally or build their own specialised images on top of
stable toolsets.

@@

\textbf{\href{http://gigantum.com/}{Gigantum}} also its own image stack
(\url{https://github.com/gigantum/base-images}). \ldots{}

@@

\textbf{\pkg{stevedore}}
(\url{https://cran.r-project.org/package=stevedore}) currently the most
powerful Docker client in R.

\textbf{\pkg{AzureContainers}}
(\url{https://cran.r-project.org/package=AzureContainers}) is an
interface to a number of container-related services in Microsoft's
\href{https://azure.microsoft.com/}{Azure Cloud}, namely
\href{https://azure.microsoft.com/en-us/services/container-instances/}{Container
Instances},
\href{https://azure.microsoft.com/en-us/services/container-registry/}{Container
Registry}, and
\href{https://azure.microsoft.com/en-us/services/kubernetes-service/}{Kubernetes
Service}. While it is mainly intended for working with Azure, as a
convenience feature it includes lightweight, cross-platform shells to
Docker and Kubernetes (tools \texttt{kubectl} and \texttt{helm}). These
can be used to create and manage arbitrary Docker images and containers,
as well as Kubernetes clusters on any platform or cloud service.

@@

R has been historically viewed as a tool for analysis and scientific
research--not for creating software that corporations can rely on to
continously run in real time. However, thanks to advancements in R
running as a web service, along with with the ability to deploy R in
Docker containers, modern enterprises are now capable of having
real-time machine learning powered by R.

The backbone of this work is the \pkg{plumber}
(\url{https://www.rplumber.io/}) package. The plumber package allows R
to be used to host web services through RESTful APIs. When R is running
on a server, a different server can ask R to make a computation through
an HTTP request and get an immediate response. For instance, if a data
scientist at a company creates a machine learning model to tell which
customers are likely to make a repeat purchase, if they serve the model
through a plumber API then someone else on the marketing team could
write software that sends different emails depending on that real-time
prediction. Plumber syntax is simple to read and understand: the API is
created by commenting above function names in an R script.

@@

@@

  \begin{itemize}
  \tightlist
  \item
    To enable quick deployments of key R services such as RStudio and
    Shiny onto cloud virtual machines (VMs), this package utilises
    Dockerfiles to move the labour of setting up those services from the
    user to a premade Docker image. For example, by specifying the
    template \texttt{template="rstudio"} in
    \texttt{gce\_vm\_template()/gce\_vm()} an up to date RStudio Service
    image is launched. Specifying \texttt{template="rstudio-gpu"} will
    launch an RStudio Server image with a GPU attached, etc.\\
  \end{itemize}
\item
  \href{https://github.com/sckott/analogsea}{\texttt{analogsea}}
  (digital ocean R client)
\item
  \href{https://www.rplumber.io/docs/hosting.html\#docker}{\texttt{plumber}}
\item
  \href{https://github.com/tmobile/r-tensorflow-api}{production
  deployment of neural networks} \citep[\citet{nolistic}]{jnolis}
\item
  Kubernetes - A popular platform for managing Docker containers is
  \href{https://kubernetes.io/}{Kubernetes}, which is used for a wide
  variety of containerized applications. It may be your organisation has
  a Kubernetes cluster already for other applications. Docker containers
  are used within Kubernetes clusters to hold native code, for which
  Kubernetes creates a framework around network connections and scaling
  of resources up and down. An introduction on using
  \href{https://code.markedmondson.me/r-on-kubernetes-serverless-shiny-r-apis-and-scheduled-scripts/}{Kubernetes
  with R is published at this blog post}.
\item
  \href{https://cran.r-project.org/package=AzureContainers}{\texttt{AzureContainers}}:
  Umbrella package for working with containers in Microsoft's Azure
  cloud. Provides interfaces to three Azure services: Container
  Registry, Container Instances (for running individual containers), and
  Kubernetes Services (for orchestrated deployments)
\end{itemize}

While Plumber provides the low-level infrastructure for turning a
statistical model into a predictive service, for production purposes it
is usually necessary to take into account considerations such as
scalability, reliability and ease of management. A single container is
limited in the volume of requests it can service. Furthermore, if the
container goes down, or is occupied with processing a computationally
intensive task, the service becomes unavailable until it is restarted.

These issues can be mitigated by deploying not a single container, but a
\emph{cluster} of containers, orchestrated as a single deployment. The
deployment system of choice at the time of writing is
\href{https://kubernetes.io}{Kubernetes}. This allows you to deploy a
single image to a pool of servers, which can then serve requests in
parallel. Client-side requests are made to the cluster ingress endpoint,
which then redirects them to individual servers. The orchestrator can
identify cluster nodes that have failed, and restart them automatically;
it can also detect when there is no capacity to service more requests,
and start additional servers and containers as needed.

While these load-balancing and autoscaling properties of a Kubernetes
cluster are invaluable, they are still limited by the hardware resources
that it has access to. A cluster cannot scale out if it has run out of
servers. This issue in turn can be addressed by deploying not to
on-premise servers but to the cloud, where effectively unlimited
computing resources are available (as long as one can pay for them).
From a management perspective, cloud deployments have the additional
advantage that hardware maintenance becomes the responsibility of the
cloud provider, eliminating another pain point. Orchestrator services
are available from all the major cloud service providers, including
Amazon Web Services (AWS), Google Cloud Platform, and Microsoft Azure.

An example of an R package for working with containers in the cloud is
AzureContainers, which interacts with three Azure services: Container
Instances (for running a single container), Container Registry (a
private docker registry service), and Kubernetes Service (for Kubernetes
on Azure). AzureContainers provides a lightweight yet powerful interface
for working with Resource Manager, the framework for deploying and
managing arbitrary resources in Azure. On the client side, it provides
simple shells to the Docker, Kubectl and Helm tools that are commonly
used for managing containers and Kubernetes clusters.

Here is some simple code that pushes a Docker image to an Azure
Container Registry, and then creates a deployment using Azure Kubernetes
Service. This is a slightly modified version of the code from the
``Deploying a prediction service with Plumber'' AzureContainers
vignette. The code makes use of a yaml file to define the deployment and
predictive service, which is standard practice with Kubernetes; the yaml
can be seen in the aforementioned vignette.

\begin{verbatim}
library(AzureContainers)
# create a resource group for our deployments
deployresgrp <- AzureRMR::get_azure_login()$
    get_subscription("subscription_id")$
    create_resource_group("deployresgrp", location="australiaeast")

### create a container registry
deployreg_svc <- deployresgrp$create_acr("deployreg")

# build our image (a random forest on the Boston housing data, available in the MASS package)
call_docker("build -t bos-rf .")

# upload the image to Azure
deployreg <- deployreg_svc$get_docker_registry()
deployreg$push("bos-rf")

### create a Kubernetes cluster with 2 nodes
deployclus_svc <- deployresgrp$create_aks("deployclus", agent_pools=aks_pools("pool1", 2))

# grant the cluster pull access to the registry
gr <- AzureGraph::get_graph_login()
aks_app_id <- deployclus$properties$servicePrincipalProfile$clientID
reg$add_role_assignment(gr$get_app(aks_app_id), "Acrpull")

# get the cluster endpoint
deployclus <- deployclus_svc$get_cluster()

# create and start the service
deployclus$create("bos-rf.yaml")
\end{verbatim}

Once the service has been created, we can check on its status and obtain
predictions:

\begin{verbatim}
deployclus$get("deployment bos-rf")
#> Kubernetes operation: get deployment bos-rf  --kubeconfig=".../kubeconfigxxxx"
#> NAME      DESIRED   CURRENT   UP-TO-DATE   AVAILABLE   AGE
#> bos-rf    1         1         1            1           5m

deployclus$get("service bos-rf-svc")
#> Kubernetes operation: get service bos-rf-svc  --kubeconfig=".../kubeconfigxxxx"
#> NAME         TYPE           CLUSTER-IP   EXTERNAL-IP     PORT(S)          AGE
#> bos-rf-svc   LoadBalancer   10.0.8.189   xxx.xxx.xxx.xxx 8000:32276/TCP   5m

response <- httr::POST("http://xxx.xxx.xxx.xxx:8000/score",
    body=list(df=MASS::Boston[1:10,]), encode="json")
httr::content(response, simplifyVector=TRUE)
#> [1] 25.9269 22.0636 34.1876 33.7737 34.8081 27.6394 21.8007 22.3577 16.7812 18.9785
\end{verbatim}

Note that Plumber, by itself, provides no authentication or security
features; anybody who knows the cluster's IP address can access the
predictive service. At minimum, a basic authentication layer should be
added to any Kubernetes deployment. Standard Kubernetes procedure
involves creating an ingress controller, for example with the nginx or
traefik reverse proxy software, and one can then add
application-specific authentication layers on top of that (for example,
to authenticate with a user's Azure Active Directory or LDAP
credentials). More details can be found from most Kubernetes learning
resources, and are outside the scope of this paper.

@@

Docker is perfectly suited for packaging and executing any software and
data, and is a good candidate for building complex processing pipelines
that are independent from the host programming language. Due to its
original use case (see~\nameref{introduction}), Docker has no standard
mechanisms for chaining containers together; it lacks definitions and
protocols for environment variables, volume mounts and/or ports that
could enable transfer of input (parameters and data) and output
(results) to and from containers. Some packages, e.g.~\pkg{containerit},
do provide Docker images that can be used similarly to CLIs, but their
usage is
cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.

\textbf{\pkg{outsider}} (\url{https://antonellilab.github.io/outsider/})
tackles the problem of integrating external programs into an R workflow
\citep{bennett_outsider_2020}. Installation and usage of external
programs can be difficult, convoluted and even impossible if the
platform is incompatible. Therefore \pkg{outsider} uses the
platform-independent Docker images to encapsulate processes in
\emph{outsider modules}. Each outsider module has a \texttt{Dockerfile}
and an R package with functions for interacting with the encapsulated
tool. Using only R functions, an end-user can install a module with the
\pkg{outsider} package and then call module code to integrate a tool
into their own workflow seamlessly. The \pkg{outsider} package and
module manage the containers and handle the transmission of arguments
and the transfer of files to and from a container. These functionalities
also allow a user to launch module code on a remote machine via SSH,
expanding the potential computational scale. Outsider modules can be
hosted code-sharing services, e.g., on GitHub, and \pkg{outsider}
contains discovery functions for them.

@@

@@

The R packages and use cases presented show the positive drive in the
community to innovate, but also the challenge of keeping up with a
continuously evolving landscape. The use cases contributed by co-authors
also have a degree of overlap, which can be expected as a common
language and understanding of good practices is still taking shape.
Also, the ease with which one can create a complex software systems,
such as an independent Docker image stack, to serve one's specific needs
leads to some fragmentation. Furthermore fragmentation may not be a bad
sign but instead a reflection of a growing market, which is able to
sustain multiple related efforts. With the maturing of core building
blocks, such as the Rocker suite of images, more systems will be built
successfully but will also be behind the curtains. Docker alone, as a
flexible core technology, is not a feasible level of collaboration and
abstraction. Instead, the use cases and applications observed in this
work provide a more useful division.

>>>>

% !TeX root = RJwrapper.tex
\title{The Rockerverse: Images, Packages, and Applications for Containerization
with R}
\author{by Daniel Nüst, Dom Bennett, Robrecht Cannoodt, Dav Clark, Gergely Daroczi, Dirk Eddelbuettel, Mark Edmondson, Colin Fay, Ellis Hughes, Sean Lopp, Ben Marwick, Heather Nolis, Jacqueline Nolis, Hong Ooi, Karthik Ram, Noam Ross, Lori Shepherd, Nitesh Turaga, Craig Willis, Nan Xiao}

@@

\begin{figure}
\centering
\includegraphics{/rockerverse/rockerverse.png}
\caption{Rockerverse hex sticker}
\end{figure}

@@

\textbf{Renjin} is an interpreter for the R language running on top of
the \href{https://en.wikipedia.org/wiki/Java_virtual_machine}{Java
Virtual Machine} (JVM) providing full two-way access between Java and R
code \citep{wikipedia_renjin_2018}. It was developed to combine the
benefits of R, such scripting and extension packages, with the JVM's
advantages in the areas of security, cross-platform availability, and
established position in enterprise settings. R extension packages need
to be specially compiled and are distributed via the Java package
manager \href{https://en.wikipedia.org/wiki/Apache_Maven}{Apache Maven},
cf. \url{http://packages.renjin.org/packages} for available packages.
Packages are loaded on demand, i.e., at the first call to
\texttt{library()}. Not all R packages, especially one linking to binary
libraries, are available, e.g., \texttt{rgdal}
\footnote{\href{http://packages.renjin.org/package/org.renjin.cran/rgdal/1.4-4/build/1}{http://packages.renjin.org/package/org.renjin.cran/rgdal/1.4-4/build/1}}.
There are no offical Docker images for Renjin, but community-maintained
images for selected releases only are available under
\texttt{nuest/renjin} on Docker~Hub and GitHub at
\url{https://hub.docker.com/r/nuest/renjin} and
\url{https://github.com/nuest/renjin-docker} respectively. These images
expose the command line interface of Renjin in a similar fashion as
Rocker images and allow an easy evaluation of Renjin's suitability, but
are not intended for production use.

@@

\textbf{FastR} (\url{https://github.com/oracle/fastr}) is ``A
high-performance implementation of the R programming language, built on
GraalVM'' \citep{oracle_labs_oraclefastr_2020}. It is developed by
Oracle, connects R to the GraalVM ecosystem
\citep{wikipedia_graalvm_2019}, and also claims superior performance but
also targets full compatibility with base R
\citep{oracle_labs_oraclefastr_2020}. There are no offical Docker images
provided, but indepedently maintained experimental images and
\texttt{Dockerfile}s are provided by
\href{https://github.com/nuest/fastr-docker}{nuest/fastr-docker}, e.g.
\href{https://cloud.docker.com/repository/docker/nuest/fastr/}{\texttt{nuest/fastr}
on Docker Hub}.

@@

The \href{https://github.com/jupyter/docker-stacks/}{\textbf{Jupyter
Docker Stacks}} project are a set of ready-to-run Docker images
containing Jupyter applications and interactive computing tools
\citep{project_jupyter_jupyter_2018}. The \texttt{jupyter/r-notebook}
image includes R and ``popular packages'', and naturally also the
IRKernel (\url{https://irkernel.github.io/}), an R kernel for Jupyter so
that Jupyter Notebooks can contain R code cells. R is also included in
the catchall \texttt{jupyter/datascience-notebook}
image\footnote{\href{https://jupyter-docker-stacks.readthedocs.io/en/latest/using/selecting.html}{https://jupyter-docker-stacks.readthedocs.io/en/latest/using/selecting.html}}.
For example, these images allow users to quickly start a Jupyter
Notebook server locally or build their own specialised images on top of
stable toolsets. R is installed using the Conda package manager, which
can manage environments for various programming languages, pinning both
the R version and the versions of R
packages\footnote{See \texttt{jupyter/datascience-notebook}'s \texttt{Dockerfile} at \href{https://github.com/jupyter/docker-stacks/blob/master/datascience-notebook/Dockerfile\#L47}{https://github.com/jupyter/docker-stacks/blob/master/datascience-notebook/Dockerfile\#L47}.}.

@@

\textbf{Gigantum} (\url{http://gigantum.com/} is a platform for open and
decentralized data science with a focus on using automation and
user-friendly tools for easy sharing of reproducible computational
workflows. Gigantum builds on the \emph{Gigantum Client} (running either
locally or on a remote server) for development and execution of
data-focused \emph{Projects}, which can be stored and shared via the
\emph{Gigantum Hub} or via a zipfile export. The Client is a user
friendly interface to a backend using Docker containers to package,
build, and run Gigantum projects. It is configured to use a default set
of Docker base images (\url{https://github.com/gigantum/base-images}),
and users are able to define and configure their own custom images. The
available images include two with R based on Ubuntu Linux have the
\href{https://launchpad.net/~marutter/+archive/ubuntu/c2d4u3.5/}{\texttt{c2d4u}
CRAN PPA} pre-configured for installation of binary R
packages\footnote{\href{https://docs.gigantum.com/docs/using-r}{https://docs.gigantum.com/docs/using-r}}.
The R images vary in the included authoring environment, i.e., Jupyter
in \texttt{r-tidyverse} or both Jupyter \& RStudio in
\texttt{rstudio-server}. The independent image stack can be traced back
to the Gigantum environment and its features. The R images are based on
Gigantum's \texttt{python3-minimal} image, originally to keep the
existing front-end configuration, but also to provide a consistent
Python-to-R interoperability. The \texttt{Dockerfile}s also use build
args to specify bases, for example for different version of NVIDIA CUDA
for GPU
processing\footnote{See \href{https://github.com/gigantum/base-images/blob/master/_templates/python3-minimal-template/Dockerfile}{https://github.com/gigantum/base-images/blob/master/\_templates/python3-minimal-template/Dockerfile} for the \texttt{Dockerfile} of \texttt{python3-minimal}.}
so that appropriate GPU drivers can be enabled automatically when
supported. Furthermore, Gigantum's focus lies on environment management
via GUI and ensuring a smooth user interaction, e.g., with reliable and
easy conflict detection and resolution. For this reason, project
repositories store authoritative package information in a separate file
per package, allowing Git to directly detect conflicts and changes. A
Dockerfile is generated from this description that inherits from the
specified base image, and additional custom Docker instructions may be
appended by users, though Gigantum's default base images do not
currently include the \texttt{littler} tool, which is used by Rocker to
install packages within \texttt{Dockerfile}s. Because of these
specifics, instructions from \texttt{rocker/r-ubuntu} could \emph{not}
be readily re-used in this image stack (see
Section~\nameref{conclusion}). Both approaches enable \texttt{apt} as an
installation method, and this is exposed via the GUI-based environment
management,\footnote{See \href{https://docs.gigantum.com/docs/environment-management}{https://docs.gigantum.com/docs/environment-management}}.
The image build and publication process is scripted with Python and JSON
template configuration files, unlike Rocker which relies on plain
\texttt{Dockerfile}s. A minor reason in the inception of the images were
also project constraints requiring a Rocker-incompatible licensing of
the \texttt{Dockerfile}s, i.e., the MIT License.

\textbf{Whole Tale} (\url{https://wholetale.org}) is a National Science
Foundation (NSF) funded project\footnote{NSF Award OAC-1541450}
developing a scalable, open-source, multi-user platform for reproducible
research \citep{brinckman2019, chard2019a}. A central goal of the
platform is to enable researchers to easily create and publish
executable research
objects\footnote{In Whole Tale a `tale` is a research object that contains metadata, data (by copy or reference), code, narrative, documentation, provenance, and information about the computational environment to support computational reproducibility}
associated with published research \citep{chard2019b}. For R users,
Whole Tale attempts to combine the strengths of the Rocker Project's
curated Docker images with the Jupyter Project's
\texttt{repo2docker}\footnote{See \href{https://repo2docker.readthedocs.io/}{https://repo2docker.readthedocs.io/}}
component to simplify the customization of R-based environments for
researchers with limited experience with either Docker or Git. Using
Whole Tale, researchers can create and publish Rocker-based reproducible
research objects to a growing number of repositories including DataONE
member nodes, Zenodo and soon Dataverse. Additionally, Whole Tale
supports automatic data citation and is working on capabilities for
image preservation and provenance capture to improve the transparency of
publish computational research artifacts
\citep{mecum2018, mcphillips2019}.

@@

\textbf{\pkg{stevedore}}
(\url{https://cran.r-project.org/package=stevedore}) is currently the
most powerful Docker client in R. It interfaces with the Docker daemon
over the Docker HTTP API\footnote{} via a Unix socket on Linux or MacOS,
over a named pipe on Windows, or over an HTTP/TCP connection. The
package is the only one not using system calls to the \texttt{docker}
CLI for managing images and containers and easily exposes connections to
remote Docker daemons, which has to be configured on the Docker level
otherwise. Using the API gives access to more information and is system
independent and likely more reliable than parsing command line output.
\pkg{stevedore}'s own interface is automatically generated based on the
OpenAPI specification of the Docker daemon, but still similar to the
Docker CLI. The interface is similar to R6 objects, in that a
\texttt{stevedore\_object} has a number of functions attached to it that
can be called, and multiple specific versions of the Docker API can be
supported thanks to the automatic
generation\footnote{See \href{https://github.com/richfitz/stevedore/tree/master/inst/spec}{https://github.com/richfitz/stevedore/tree/master/inst/spec} and \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

\textbf{\pkg{AzureContainers}}
(\url{https://cran.r-project.org/package=AzureContainers}) is an
interface to a number of container-related services in Microsoft's
\href{https://azure.microsoft.com/}{Azure Cloud}, see
Section~\nameref{azurecontainers}. While it is mainly intended for
working with Azure, as a convenience feature it includes lightweight,
cross-platform shells to Docker and Kubernetes (tools \texttt{kubectl}
and \texttt{helm}). These can be used to create and manage arbitrary
Docker images and containers, as well as Kubernetes clusters on any
platform or cloud service.

@@

R has been historically viewed as a tool for analysis and scientific
research--not for creating software that corporations can rely on to
continously run in real time. However, thanks to advancements in R
running as a web service, along with with the ability to deploy R in
Docker containers, modern enterprises are now capable of having
real-time machine learning powered by R. One of the first
implementations of having R respond to client reqests were TCP/IP and
local socket servers, such as \CRANpkg{Rserve} and \CRANpkg{svSocket},
which allowed other local or remote processes to make binary requests to
interact with R without the need to link to R at compile time and
connected other programming languages with R. R could also empower web
services since at least 2007, when
\href{http://www.rapache.net}{rApache} was presented at the useR!
conference \citep{horner_rapache_2013}. rApache runs an R interpreter
inside the Apache web server, which makes it possible and very efficient
to build scalable R-based APIs and web services.

Such rApache applications have been used at
\href{http://rapporter.net}{rapporter.net}, transforming parametrized R
Markdown-like documents into interactive web applications for reporting.
A similar design was implemented at
\href{https://www.card.com/}{CARD.com}, where the optimization of the
acquisition flow and the real-time analysis of debit card transactions
both relied on an efficient combination of Rserve and rApache running in
Docker containers: the R session behind Rserve acted as a read-only,
exteremely fast and scalabale, in-memory database for the many
concurrent rApache processes responding to the live-scoring requests of
various divisions of the company. Similar dockerized R scripts were
responsible for the ETL processes and even the client-facing email, text
message and push notification alerts sent in real-time based on card
transactions. The related Docker images were made available at
\url{https://github.com/cardcorp/card-rocker}. The images extend
\texttt{rocker/r-base}, e.g., with an SSH client and a workaround for
being able to mount SSH keys from the host, Pandoc, the Amazon Web
Services (AWS) SDK, and Java, which is required by the AWS SDK. The AWS
SDK allowed to run R consumers reading from real-time data processing
streams of \href{https://aws.amazon.com/kinesis/}{AWS Kinesis}. The
applications were deployed on Amazon Elastic Container Service
(\href{https://aws.amazon.com/ecs/}{ECS}). The main learnings from using
R in Docker was the imporance of not only pinning the R package versions
via MRAN, but also moving away from Debian testing to an LTS
distribution to minimize the risk of breaking the automated builds of
the Docker images and production jobs.

A more recent related work is the \pkg{plumber}
(\url{https://www.rplumber.io/}) package, which makes it even easier for
R to be used to host web services through RESTful APIs. When R is
running on a server, a different server can ask R to make a computation
through an HTTP request and get an immediate response. For instance, if
a data scientist at a company creates a machine learning model to tell
which customers are likely to make a repeat purchase, if they serve the
model through a plumber API then someone else on the marketing team
could write software that sends different emails depending on that
real-time prediction. Plumber syntax is simple to read and understand:
the API is created by commenting above function names in an R script.

@@

\href{http://restrserve.org}{RestRserve} is an alternative R web API
framework with the main focus on being high-performant and robust. It
also ships automated Docker builds at
\url{https://hub.docker.com/r/rexyai/restrserve/dockerfile}.

@@

  \begin{itemize}
  \tightlist
  \item
    To enable quick deployments of key R services such as RStudio and
    Shiny onto cloud virtual machines (VMs), this package utilises
    Dockerfiles to move the labour of setting up those services from the
    user to a premade Docker image. For example, by specifying the
    template \texttt{template="rstudio"} in
    \texttt{gce\_vm\_template()/gce\_vm()} an up to date RStudio Service
    image is launched. Specifying \texttt{template="rstudio-gpu"} will
    launch an RStudio Server image with a GPU attached, etc.\\
  \end{itemize}
\item
  \href{https://github.com/sckott/analogsea}{\texttt{analogsea}}
  (digital ocean R client)
\item
  \href{https://www.rplumber.io/docs/hosting.html\#docker}{\texttt{plumber}}
\item
  \href{https://github.com/tmobile/r-tensorflow-api}{production
  deployment of neural networks} \citep[\citet{nolistic}]{jnolis}
\item
  Kubernetes - A popular platform for managing Docker containers is
  \href{https://kubernetes.io/}{Kubernetes}, which is used for a wide
  variety of containerized applications. It may be your organisation has
  a Kubernetes cluster already for other applications. Docker containers
  are used within Kubernetes clusters to hold native code, for which
  Kubernetes creates a framework around network connections and scaling
  of resources up and down. An introduction on using
  \href{https://code.markedmondson.me/r-on-kubernetes-serverless-shiny-r-apis-and-scheduled-scripts/}{Kubernetes
  with R is published at this blog post}.
\end{itemize}

\label{azurecontainers} \textbf{\pkg{AzureContainers}} is an umbrealla
package providing interfaces the three services of Microsoft's Azure
Cloud, namely
\href{https://azure.microsoft.com/en-us/services/container-instances/}{Container
Instances} for running individual containers,
\href{https://azure.microsoft.com/en-us/services/container-registry/}{Container
Registry} for private image distribution, and
\href{https://azure.microsoft.com/en-us/services/kubernetes-service/}{Kubernetes
Service} for orchestrated deployments. While a package like
\pkg{plumber} provides the infrastructure for turning an R workflow a
service, for production purposes it is usually necessary to take into
account scalability, reliability and ease of management. A single
container is limited in the volume of requests it can service, and if
the container goes down or is occupied with processing a computationally
intensive task, the service becomes unavailable. These issues can be
mitigated by deploying not a single container, but a \emph{cluster} of
containers, orchestrated as a single deployment, e.g., with
\href{https://kubernetes.io}{Kubernetes}. A cluster features
load-balancing, autoscaling of containers across numerous servers (in
the cloud or on premise), and restarting failed ones.
\pkg{AzureContainers} provides a lightweight yet powerful interface for
working with Resource Manager, the framework for deploying and managing
arbitrary resources in Azure. On the client side, it provides simple
shells to the Docker, and to Kubernetes' \texttt{kubectl} and
\texttt{helm} tools, that are needed for managing Kubernetes clusters.
The package simplifies complex infrastructure management to a number of
R funcction calls, given an Azure account with sufficient
credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.

@@

Docker is perfectly suited for packaging and executing any software and
data, and is a good candidate for building complex processing pipelines
that are independent from the host programming language. Due to its
original use case (see~\nameref{introduction}), Docker has no standard
mechanisms for chaining containers together; it lacks definitions and
protocols for environment variables, volume mounts and/or ports that
could enable transfer of input (parameters and data) and output
(results) to and from containers. Some packages, e.g. \pkg{containerit},
do provide Docker images that can be used similarly to CLIs, but their
usage is
cumbersome\footnote{\href{https://o2r.info/containerit/articles/container.html}{https://o2r.info/containerit/articles/container.html}}.

\textbf{\pkg{outsider}} (\url{https://antonellilab.github.io/outsider/})
tackles the problem of integrating external programs into an R workflow
\citep{bennett_outsider_2020}. Installation and usage of external
programs can be difficult, convoluted and even impossible if the
platform is incompatible. Therefore \pkg{outsider} uses the
platform-independent Docker images to encapsulate processes in
\emph{outsider modules}. Each outsider module has a \texttt{Dockerfile}
and an R package with functions for interacting with the encapsulated
tool. Using only R functions, an end-user can install a module with the
\pkg{outsider} package and then call module code to integrate a tool
into their own R-based workflow seamlessly. The \pkg{outsider} package
and module manage the containers and handle the transmission of
arguments and the transfer of files to and from a container. These
functionalities also allow a user to launch module code on a remote
machine via SSH, expanding the potential computational scale. Outsider
modules can be hosted code-sharing services, e.g., on GitHub, and
\pkg{outsider} contains discovery functions for them.

@@

\hypertarget{publication-reproducibility-audits}{%
\subsection{Publication Reproducibility
Audits}\label{publication-reproducibility-audits}}

A recommendation of the recent National Academies' report on
\emph{Reproducibility and Replicability in Science} is that journals
``consider ways to ensure computational reproducibility for publications
that make claims based on computations'' \citep{NASEM2019}. In fields
such as Political Science and Economics, journals are increasingly
adopting policies requiring authors to publish the code and data
required to reproduce computational findings reported in published
manuscripts, subject to independent verification
\citep{Jacoby2017,Vilhuber2019,Alvarez2018,Christian2018,Eubank2016,King1995}.
Problems with the computational environment, installation and
availability of software dependencies are common. R is gaining
popularity in these communities (in a sample of 105 replication packages
published by the \emph{American Journal of Political Science} (AJPS)
over 65\% use R). Projects such as NSF Whole Tale use the Rocker Project
community images with the goal of improving the reproducibility of
published research artifacts and simplifying the publication and
verification process for both authors and journals by reducing errors
and time spent specifying the environment.

@@

The R packages and use cases presented show the positive drive in the
community to innovate, but also the challenge of keeping up with a
continuously evolving landscape. The use cases contributed by co-authors
also have a degree of overlap, which can be expected as a common
language and understanding of good practices is still taking shape.
Also, the ease with which one can create a complex software systems with
Docker, such as an independent Docker image stack, to serve one's
specific needs leads to parallel developments. This ease-of-DIY in
combination with the difficulty to compose environments based on
\texttt{Dockerfile}s alone is a further reason for
\textbf{fragmentation}. Instructions can be outsourced into
distributable scripts and then copied into the image during build, but
that make \texttt{Dockerfile} hard to read and adds a layor of
complexity. Despite the different image stacks presented here, the
pervasiveness of Rocker can be traced back to its maintainers and the
user community valueing collaboration and shared starting points over
impulses to create individual solutions. Aside from that, fragmentation
may not be a bad sign, but instead a reflection of a growing market,
which is able to sustain multiple related efforts. With the maturing of
core building blocks, such as the Rocker suite of images, more systems
will be built successfully but will also be behind the curtains. Docker
alone, as a flexible core technology, is not a feasible level of
collaboration and abstraction. Instead, the use cases and applications
observed in this work provide a more useful division.

Nonetheless, at least on the level of R packages some
\textbf{consolidation} seems in order, e.g., to reduce the number of
packages creating \texttt{Dockerfile}s from R code or controlling the
Docker daemon with R code. It remains to be seen which approach to
control Docker, via the Docker API as \texttt{stevedore} or via system
calls as \texttt{dockyard}/\texttt{docker}/\texttt{dockr}, is more
sustainable, or if the question will be answered by the endurance of
maintainers and sufficient funding. Similarly, the capturing of
environments and their serialization in form of a \texttt{Dockerfile}
currently happens at different levels of abstraction and re-use of
functionality seems reasonable, e.g. \texttt{liftr} could generate the
environment with \texttt{containerit}, which in turn may use
\texttt{dockerfiler} for low level objects representing a
\texttt{Dockerfile}. In this consolidation, the Rocker~Project could
play the role of coordinating entity. Though for the moment, the sign of
the times points to more experimentation and feature growth, e.g.~images
for GPU-based computing and artificial intelligence. Even with coding
being more and more accepted as a required, and achievable skill, an
easier access, for example by exposing containerisation benefits via
simple user interfaces in the users' IDE, could be an important next
step. Currently containerisation happens more in the background at the
system level.
